\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-05-21 doc/USER_MANUAL/06_fault_sources.tex
Buildbot ran the code beautifier.

Here is an example CUBIT script to generate a mesh with split nodes
for a buried vertical strike-slip fault:
\begin{lyxcode}
~reset~~\\
~brick~x~10~y~10~z~10~~\\
~webcut~volume~all~with~plane~xplane~~\\
~webcut~volume~all~with~plane~yplane~~\\
~webcut~volume~all~with~plane~xplane~offset~3~~\\
~webcut~volume~all~with~plane~zplane~offset~3~~\\
~webcut~volume~all~with~plane~zplane~offset~-3~~\\
~imprint~all~~\\
~merge~all~~\\
~unmerge~surf~160~~\\
~mesh~vol~all~~\\
~set~node~constraint~off~~\\
~node~in~surf~168~move~X~0~Y~0.01~Z~0~~\\
~node~in~surf~160~move~X~0~Y~-0.01~Z~0~\\

\end{lyxcode}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.3]{figures/faultmesh.jpg} \\
\includegraphics[scale=0.3]{figures/surf168.jpg} 
\includegraphics[scale=0.3]{figures/splitnodes.jpg}
\par\end{centering}

\caption{Screenshots of the CUBIT example for the embedded fault with split
nodes: The entire mesh is decomposed into several volumes bounding
the fault (top), the fault surface 168 is shown as orange squares
(middle) and split nodes of the fault shown as orange dots(bottom).
Note that only interior nodes of the fault are split while those on
the edges of the fault surface are touching each other.}

\label{fig:examples.splitnodes}
\end{figure}

\begin{lyxcode}

\end{lyxcode}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace.pdf} \\

@@

Note that you should avoid gaps in the list of indices of mesh objects
with the following CUBIT command:
\begin{lyxcode}
$\mathtt{compress\: ids\: hex\: face\: edge\: node}$
\end{lyxcode}
(otherwise you will get a segmentation fault during domain decomposition)

@@

The mesh generated in CUBIT needs to be processed and exported in
a format compatible with SPECFEM3D. This is achieved in the Python
scripts by calling the Python-CUBIT interface functions defined in
the CUBIT directory:
\begin{enumerate}
\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
must be called to set up the absorbing boundaries database
\item Function $\mathtt{fault\_input}$ must be called once for each fault
to set up the fault database
\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
at the very end of the script to export the mesh in a SPECFEM3D format.
\end{enumerate}
The functions in \#1 and \#3 are for the bulk and are documented in
Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
\begin{description}
\item [{Function:}] $\mathtt{fault\_input}$
\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
file
\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
\item [{Inputs:}]~\end{description}
\begin{itemize}
\item id\_fault integer index assigned to the fault. (The user must number
all the faults, starting at 1, with unit increments)
\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
the fault.
\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
the fault. (The user must decide which side of the fault is side 1.
This choice affects the sign conventions of fault quantities as explained
in Section \ref{sec:Sign-Convention-for}).\end{itemize}
\begin{description}
\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
fault id (id\_fault).
\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:\end{description}
\begin{lyxcode}
$\hphantom{}$A1~=~{[}168{]}

$\hphantom{}$A2~=~{[}160{]}

$\hphantom{}$fault\_input(1,A1,A2)
\end{lyxcode}

>>>>

Here is an example CUBIT script to generate a mesh with split nodes
for a buried vertical strike-slip fault:
{\footnotesize
\begin{verbatim}
 reset
 brick x 10 y 10 z 10
 webcut volume all with plane xplane
 webcut volume all with plane yplane
 webcut volume all with plane xplane offset 3
 webcut volume all with plane zplane offset 3
 webcut volume all with plane zplane offset -3
 imprint all
 merge all
 unmerge surf 160
 mesh vol all
 set node constraint off
 node in surf 168 move X 0 Y 0.01 Z 0
 node in surf 160 move X 0 Y -0.01 Z 0
\end{verbatim}
}
%
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.3]{figures/faultmesh.jpg} \\
\includegraphics[scale=0.3]{figures/surf168.jpg}
\includegraphics[scale=0.3]{figures/splitnodes.jpg}
\par\end{centering}

\caption{Screenshots of the CUBIT example for the embedded fault with split
nodes: The entire mesh is decomposed into several volumes bounding
the fault (top), the fault surface 168 is shown as orange squares
(middle) and split nodes of the fault shown as orange dots(bottom).
Note that only interior nodes of the fault are split while those on
the edges of the fault surface are touching each other.}
\label{fig:examples.splitnodes}
\end{figure}

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.55]{figures/splitnodes_surfacetrace.pdf} \\

@@

Note that you should avoid gaps in the list of indices of mesh objects
with the following CUBIT command:
{\footnotesize
\begin{verbatim}
compress ids hex face edge node
\end{verbatim}
}
(otherwise you will get a segmentation fault during domain decomposition)

@@

The mesh generated in CUBIT needs to be processed and exported in
a format compatible with SPECFEM3D. This is achieved in the Python
scripts by calling the Python-CUBIT interface functions defined in
the CUBIT directory:
\begin{enumerate}
\item Function $\mathtt{define\_bc}$ (or $\mathtt{boundary\_definition.py}$)
must be called to set up the absorbing boundaries database
\item Function $\mathtt{fault\_input}$ must be called once for each fault
to set up the fault database
\item Function $\mathtt{cubit2specfem3d.export2SPECFEM3D}$ must be called
at the very end of the script to export the mesh in a SPECFEM3D format.
\end{enumerate}
The functions in \#1 and \#3 are for the bulk and are documented in
Section \ref{subsec:Exporting-the-Mesh}. We focus here on \#2:
\begin{description}
\item [{Function:}] $\mathtt{fault\_input}$
\item [{Purpose:}] export a fault mesh database from CUBIT to a SPECFEM3D-compliant
file
\item [{Syntax:}] fault\_input(id\_fault, ids\_surf\_1, ids\_surf\_2)
\item [{Inputs:}]~\end{description}
\begin{itemize}
\item id\_fault integer index assigned to the fault. (The user must number
all the faults, starting at 1, with unit increments)
\item ids\_surf\_1 list of CUBIT ids of all surfaces that form side 1 of
the fault.
\item ids\_surf\_2 list of CUBIT ids of all surfaces that form side 2 of
the fault. (The user must decide which side of the fault is side 1.
This choice affects the sign conventions of fault quantities as explained
in Section \ref{sec:Sign-Convention-for}).\end{itemize}
\begin{description}
\item [{Outputs:}] file $\mathtt{fault\_file\_X.dat}$, where X is the
fault id (id\_fault).
\item [{Example:}] For the example in Section \ref{sec:Mesh-Generation-with}:
\begin{verbatim}
  A1 = [168]
  A2 = [160]
  fault_input(1,A1,A2)
\end{verbatim}
\end{description}
